# Tests

::: {.callout-warning style="color: orange;"}
###### Warning

If you want to exclude this file, remove it from the mentioned chapters
in `_quarto.yml`.
:::

## Code Annotation with echo: fenced

```{r}
#| label: test-code-annotation
#| echo: fenced
1 + 1 # <1>
2 + 2 # <2>
```

1.  Simple
2.  Very simple

## Cross Reference to a Tibble

Cross references to the internal printing mode of tibbles do not work.
Maybe there are option you can set up with {**pillar**}?

To cross reference the output of a tibble you need to convert it to a
table.

```{r}
#| label: tbl-mytable
#| tbl-cap: "This is a test"

library(knitr)
myTable <- tibble::tibble(toss = c("w", "l", "w", "w", "w", "l", "w", "l", "w"))
kable(myTable)
```

See @tbl-mytable

## Table Width with kableExtra

```{r}
#| label: table-width

library(kableExtra)

dt <- mtcars[1:5, 1:3]

dt %>%
  kbl(caption = "Recreating booktabs style table") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```

## Cross Reference to Code Listing (not working)

```{r}
#| lst-label: lst-code1
#| lst-cap: "Some R code 1"

a <- 1
a + 1
```

I'd like to be able to reference the code chunk @lst-code1.

## Cross Reference to Code Listing (working!)

``` {#lst-code .r lst-cap="Some R code"}
a <- 1
a + 1
```

I'd like to be able to reference the code chunk @lst-code2.

```{r}
#| echo: true
#| eval: true
#| label: test-chunk
#| attr-source: '#lst-summarize lst-cap="Sum of five and five"'
5 + 5
```

You refer to the listing of this code chunk with @list-summarize.

## Annotations with Cross References

```{r}

1 + 1 # <1>
2 + 2 # <2>
3 + 3 # <3>
4 + 4 # <4>
5 + 5 # <5>
6 + 6 # <6>
```

1.  My first annotation. More in @sec-annotation-1
2.  My second annotation. More in @sec-annotation-2
3.  My third annotation. More in @sec-annotation-3
4.  My fourth annotation. More in @sec-annotation-4
5.  My fifth annotation. More in @sec-annotation-5
6.  My sixth annotation. More in @sec-annotation-6

### Annotation \<1\> {#sec-annotation-1}

More explaining text.

### Annotation \<2\> {#sec-annotation-2}

More explaining text.

### Annotation \<3\> {#sec-annotation-3}

More explaining text.

### Annotation \<4\> {#sec-annotation-4}

More explaining text.

### Annotation \<5\> {#sec-annotation-5}

More explaining text.

### Annotation \<6\> {#sec-annotation-6}

More explaining text.
